A Possibilistic Valid-time Model

### 2011-11-17 trunk/ipmu2012/preliminaries.tex
I modified the part about possibilistic variables because I needed the given information to explain the evaluation of sets...

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ underlying the possibility measure $\Pi$ is then a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

@@

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The possibility distribution $\pi_X$ gives the available knowledge about the value that $X$ takes. For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.

It is important to understand the difference between the following two concepts:
\begin{itemize}
\item
A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
\item
An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
\end{itemize}

Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\forall (u,v)\in\mathbb{R}^2:&\\
\nonumber
\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R}:& \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number. In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.

The most convenient representation for the membership function of a fuzzy number is a triangular function. The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization. Three values represent a triangular function (see Figure~\ref{fig:triangular}):
\begin{itemize}
\item
$D$ is the singleton value in the core of $\mu_M(x)$.
\item
$D-a$ is the lower value in the support of $\mu_M(x)$. 
\item
$D+b$ is the upper value in the support of $\mu_M(x)$.
\end{itemize}
Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/triangular.eps}
  \caption{Triangular function.}
  \label{fig:triangular}
\end{figure}

The most simple representation for the membership function of a fuzzy interval is a trapezoidal function. This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized. Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:

\begin{align}
\mu_T(x)
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
  \caption{Trapezoidal distribution.}
  \label{fig:trapezoidal}
\end{figure}

\subsubsection{Set evaluation by ill-known constraints}

\subsubsection{Interval evaluation by ill-known constraints}
The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.

@@

\subsection{Temporal databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema. A \textbf{chronon} is the shortest duration of time supported by the database. The time can be represented either as points or intervals \cite{655777}. Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known. There are proposals for the  fuzzyfication of the time point and the fuzzyfication of the time interval.\\
Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.

>>>>

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

@@

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
A \emph{possibilistic variable} can intuitively be seen as a variable for which the possible values are described by its possibility distribution\cite{Pon11}.

\begin{definition}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
\end{definition}

The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work\cite{Dubois88b}.

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set. In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.

In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values. The correspondences and transitions between these two representations are part of the authors current research.

%It is important to understand the difference between the following two concepts:
%\begin{itemize}
%\item
%A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
%\item
%An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
%\end{itemize}

%Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention. This possibility and necessity is noted here as:
\begin{align}
Pos(p) & = \pi_p(T) \label{propholdsposs} \\
Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
  \caption{Trapezoidal membership function}
  \label{fig:trapezoidal}
\end{figure}

The most convenient form of the membership function of a fuzzy number is a triangular function. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. A visualization is given in figure \ref{fig:triangular}. Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number.
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}
In this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/triangular.eps}
  \caption{Triangular membership function.}
  \label{fig:triangular}
\end{figure}

%\subsubsection{Set evaluation by ill-known constraints}

\subsubsection{Interval Evaluation by Ill-known Constraints}
The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.

@@

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema. A \textbf{chronon} is the shortest duration of time supported by the database. The time can be represented either as points or intervals \cite{655777}. Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known. There are proposals for the  fuzzyfication of the time point and the fuzzyfication of the time interval.\\
Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.
